---
title: 'Brooklyn: Eviction Rates Over Time'
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: embed
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(readxl)
library(tidyverse)
library(plotly)
library(tidyverse)
library(readxl)
library(plotly)
library(rgdal)
library(maps)
library(devtools)
library(leaflet)
library(maptools)
library(BAMMtools)
library(patchwork)

# Data for spaghetti plot

neighborhood_df = 
  read_excel("./data/NYC_neighborhoods_by_censustract_2010.xlsx", 
             skip = 5, 
             col_names = c("borough", "county_code", "boro_code", "census_tract", "PUMA", "nta_code", "neighborhood")
             ) %>% 
#creating 11 digit FIPS code by pasting country code (36) with `county_code` and `census_tract`
  mutate(geo_id = paste0("36", county_code, census_tract)
  )

brooklyn_data = read.csv('./data/EvictionData_NY.csv') %>% 
  janitor::clean_names() %>% 
  mutate(geoid = as.character(geoid)) %>% 
  filter(parent_location == "Kings County, New York") %>% 
  left_join(neighborhood_df, by = c("geoid" = "geo_id"))

brooklyn_data = brooklyn_data %>% 
  group_by(neighborhood, year) %>% 
  mutate(
    neighb_pop = sum(population),
    tract_weight = population/neighb_pop,
    eviction_discrepancy = eviction_filing_rate - eviction_rate,
    weighted_eviction_discrep = eviction_discrepancy*tract_weight,
    weighted_eviction_rate = eviction_rate*tract_weight,
    pct_nonwhite = 100 - pct_white,
    weighted_pct_nonwhite = pct_nonwhite*tract_weight,
    weighted_povertyrate = poverty_rate*tract_weight)
 

brooklyn_data = brooklyn_data %>% 
  summarize(
    Eviction_Rate = as.numeric(sum(weighted_eviction_rate)),
    wt_eviction_rate_discrep = sum(weighted_eviction_discrep),
    wt_pct_nonwhite = sum(weighted_pct_nonwhite),
    wt_povertyrate = sum(weighted_povertyrate)) %>% 
  ungroup() %>% 
  mutate(neighborhood = as.factor(neighborhood))

brooklyn_data <- within(brooklyn_data, Eviction_Rate[neighborhood == 'park-cemetery-etc-Brooklyn'] <- NA)
brooklyn_data <- within(brooklyn_data, wt_eviction_rate_discrep[neighborhood == 'park-cemetery-etc-Brooklyn'] <- NA)




# Data for map

eviction = 
  read.csv('./data/EvictionData_NY.csv') %>% 
  janitor::clean_names() %>% 
  filter(parent_location %in% c("New York County, New York", "Queens County, New York", "Kings County, New York", "Bronx County, New York", "Richmond County, New York"),
         population != 0) %>% 
  
  ## selecting and renaming variables
  mutate(pct_nonwhite = (100 - pct_white),
         geoid = as.character(geoid),
         year = as.ordered(year)) %>% 
  select(geoid, year, ## contextual
         evictions, eviction_filings, eviction_rate, eviction_filing_rate, renter_occupied_households, ## outcomes
         pct_nonwhite_evictiondata = pct_nonwhite, ## hypothesized predictors
         poverty_rate, rent_burden, pct_af_am, pct_renter_occupied, median_gross_rent, median_household_income, median_property_value) %>% ## other predictors
  filter(renter_occupied_households > 10) # census tracts with few renter households were creating eviction rates of 150, 200, etc.

## creating key for geoid <--> boro
geoid_boro_key = 
  read.csv('./data/EvictionData_NY.csv') %>% 
  janitor::clean_names() %>% 
  filter(parent_location %in% c("New York County, New York", "Queens County, New York", "Kings County, New York", "Bronx County, New York", "Richmond County, New York"),
         year %in% c(2010:2016)) %>% 
  select(geoid, county = parent_location) %>% 
  mutate(geoid = as.character(geoid),
         boro = recode(county, "New York County, New York" = "Manhattan", 
                               "Queens County, New York" = "Queens", 
                               "Kings County, New York" = "Brooklyn", 
                               "Bronx County, New York" = "Bronx", 
                               "Richmond County, New York" = "Staten Island")) %>% 
  distinct()

joined_data = 
  eviction %>% 
  inner_join(., geoid_boro_key, by = "geoid") 

## just brooklyn
joined_data_bklyn =
  joined_data %>% 
  filter(boro == "Brooklyn")

# Loading census tract shapefiles from NYC OpenData:
census_tracts = readOGR(dsn = "maps/2010-census-tracts.shp", encoding = "UTF-8", verbose=FALSE) 

census_tracts = 
  census_tracts[census_tracts$boro_name == "Brooklyn", ]

writeOGR(census_tracts,"./mapping_files","census_temp", driver = "ESRI Shapefile", overwrite_layer = TRUE)
census_tracts = readOGR("./mapping_files","census_temp", verbose=FALSE)

# Constructing geoid for merging
census_tracts@data = 
  census_tracts@data %>% 
  mutate(
    county_code = as.character(recode(boro_code, "1" = "061", "2" = "005", "3" = "047", "4" = "081", "5" = "085")),
    ct2010 = as.character(ct2010),
    fips_11dig = paste0("36", county_code, ct2010),
    long = coordinates(spTransform(census_tracts, CRS("+proj=longlat +datum=WGS84")))[,1],
    lat = coordinates(spTransform(census_tracts, CRS("+proj=longlat +datum=WGS84")))[,2]
    ) %>%
  unnest(c(long, lat))


```


Neighborhoods Over Time
===================================== 

Column {.sidebar}
-----------------------------------------------------------------------


```{r for_sidebar}

# Pull list of brooklyn neighborhoods from brooklyn eviction data
neighborhoods = brooklyn_data %>% distinct(neighborhood) %>% pull()


# Create checkbox input

checkboxGroupInput("neighborhood_choice", label=h3("Select Neighborhood"),
choices = neighborhoods, selected = "Williamsburg")

```

Column {.tabset}
-----------------------------------------------------------------------

### Brooklyn 2000-2016

```{r spaghetti_map}


# Spaghetti plot over time

renderPlotly({ 

brooklyn_data = 
  brooklyn_data %>%
    filter(
      neighborhood == input[["neighborhood_choice"]]
    ) 

brooklyn_data %>% 
    group_by(year) %>%
    ungroup() %>% 
    plot_ly(x=~year, y=~Eviction_Rate, type="scatter", mode="lines+markers", color = ~neighborhood) %>% 
    layout(title="Eviction Rate (%) Over Time (2000-2016) by Brooklyn Neighborhood")

})


```

### Description

The line plot illustrates the eviction rates from 2000 to 2016 across 51 distinct Brooklyn neighborhoods. Users can compare neighborhoods side-by-side and visualize differences in how this parameter has changed over time throughout Brooklyn.

**Not all neighborhoods have complete eviction data for each year.**

Map Over Time: Census Tract
===================================== 

Column {.sidebar}
-----------------------------------------------------------------------

```{r for_sidebar_2}
# Create slider widget

sliderInput("years", "Year of Observation",
    min = 2000, max = 2016, step=1, value = 2000
  )
```

Column {.tabset}
-----------------------------------------------------------------------

### Eviction Rate Over Time Map

```{r map_eviction}

# plot map


renderLeaflet({
  
  joined_data_bklyn_year = 
  joined_data_bklyn  %>%
  filter(year==input[["years"]])

# Joining data and shapefiles
census_tracts@data = left_join(census_tracts@data, joined_data_bklyn_year, by = c("fips_11dig" = "geoid"))
  
# spatial transform
census_tracts_crs = spTransform(census_tracts, CRS("+init=epsg:4326"))

# get jenks natural breaks
getJenksBreaks(census_tracts_crs$eviction_rate, 5)

#set bins based on natural breaks
eviction_rate_bins = c(0, 1.23, 2.51, 4.12, 7.62)

#set color palette
eviction_palette = colorBin('Blues', bins = eviction_rate_bins, na.color = '#d9d9d9')

leaflet::leaflet(data = census_tracts_crs) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addPolygons(
    fillColor = ~eviction_palette(eviction_rate),
    fillOpacity = 0.8,
    color = "BDBDC3",
    weight = 1,
    highlightOptions = highlightOptions(color = "black", weight = 2, bringToFront = TRUE)) %>% 
      addLegend('bottomleft',
                pal = eviction_palette,
                values = ~eviction_rate,
                title = 'Eviction Rates in NYC by census tract',
                opacity = 1)

  })

```

### Description

This map displays the census-tract level eviction rates across Brooklyn from the years 2000 to 2016. 


