---
title: "What Predicts Eviction Rates?"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
---

<style type="text/css">

h1.title {
  text-align: center;
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(plotly)
library(rgdal)
library(maps)
library(devtools)
library(leaflet)
library(maptools)
library(BAMMtools)
library(patchwork)
```

### Mapping the hypothesized predictors of eviction rates

In order to inform model building, we first observed simple visualizations of our four predictors (percentage of ESL speakers, percentage of non-White residents, rent burden, and population density) over a choropleth of eviction rates in Brooklyn by census tract in 2010.

```{r loading-data, include = FALSE}
# Loading in Brooklyn data
joined_data_bklyn_2010 = 
  readRDS('./data/joined_data_bklyn') %>%
  filter(year == "2010")

# Loading census tract shapefiles from NYC OpenData:
census_tracts = readOGR(dsn = "maps/2010-census-tracts.shp", encoding = "UTF-8") 

census_tracts = 
  census_tracts[census_tracts$boro_name == "Brooklyn", ]

# The following step removes null values from the dataset based on filtering values above
writeOGR(census_tracts,"./mapping_files","census_temp", driver = "ESRI Shapefile", overwrite_layer = TRUE)
census_tracts = readOGR("./mapping_files","census_temp")

# Constructing geoid for merging
census_tracts@data = 
  census_tracts@data %>% 
  mutate(
    county_code = as.character(recode(boro_code, "1" = "061", "2" = "005", "3" = "047", "4" = "081", "5" = "085")),
    ct2010 = as.character(ct2010),
    fips_11dig = paste0("36", county_code, ct2010),
    long = coordinates(spTransform(census_tracts, CRS("+proj=longlat +datum=WGS84")))[,1],
    lat = coordinates(spTransform(census_tracts, CRS("+proj=longlat +datum=WGS84")))[,2]
    ) %>%
  unnest(c(long, lat))

# Joining data and shapefiles
census_tracts@data = left_join(census_tracts@data, joined_data_bklyn_2010, by = c("fips_11dig" = "geoid"))

# Create map projection ready for mapping
census_tracts_crs = 
  spTransform(census_tracts, CRS("+init=epsg:4326"))
```

```{r map-aesthetics, echo=FALSE, include = FALSE}
# Creating mapping aesthetics

# Labels
esl_popup = paste0(
  "<strong>Census tract: </strong>",
  census_tracts_crs$ctlabel,
  "<br><strong>Eviction rate: </strong>",
  census_tracts_crs$eviction_rate,
  "<br><strong>% ESL: </strong>",
  census_tracts_crs$pct_eng
)

nonwhite_popup = paste0(
  "<strong>Census tract: </strong>",
  census_tracts_crs$ctlabel,
  "<br><strong>Eviction rate: </strong>",
  census_tracts_crs$eviction_rate,
  "<br><strong>% non-White: </strong>",
  census_tracts_crs$pct_nonwhite_racedata
)

rentburden_popup = paste0(
  "<strong>Census tract: </strong>",
  census_tracts_crs$ctlabel,
  "<br><strong>Eviction rate: </strong>",
  census_tracts_crs$eviction_rate,
  "<br><strong>Rent burden: </strong>",
  census_tracts_crs$rent_burden
)

popdensity_popup = paste0(
  "<strong>Census tract: </strong>",
  census_tracts_crs$ctlabel,
  "<br><strong>Eviction rate: </strong>",
  census_tracts_crs$eviction_rate,
  "<br><strong>Population Density: </strong>",
  census_tracts_crs$total_pop_densitydata
)

# Obtaining Jenks natural break points for continuous predictors
getJenksBreaks(census_tracts_crs$pct_eng, 5)
getJenksBreaks(census_tracts_crs$pct_nonwhite_racedata, 5)
getJenksBreaks(census_tracts_crs$rent_burden, 5)
getJenksBreaks(census_tracts_crs$total_pop_densitydata, 5)

# Setting bins based on Jenks natural breaks
pct_eng_bins = c(0, 14.6, 29.8, 46.9, 82.5)
nonwhite_bins = c(0, 22.9, 49.6, 77.2, 100)
rentburden_bins = c(13.9, 28.4, 34.6, 42.5, 50.0)
popdensity_bins = c(22, 2432, 3636, 5124, 8475)

# Setting color based on bins
eng_palette = colorBin('Blues', bins = pct_eng_bins, na.color = '#d9d9d9')
nonwhite_palette = colorBin('Blues', bins = nonwhite_bins, na.color = '#d9d9d9')
rentburden_palette = colorBin('Blues', bins = rentburden_bins, na.color = '#d9d9d9')
popdensity_palette = colorBin('Blues', bins = popdensity_bins, na.color = '#d9d9d9')

# Create custom legend for eviction circle marker sizes
addLegendCustom <- function(map, position, colors, labels, sizes, opacity = 0.4, title = "Eviction Rate"){
  
    colorAdditions <- paste0(colors, "; border-radius: 50%; width:", sizes, "px; height:", sizes, "px")
    labelAdditions <- paste0("<div style='display: inline-block;height: ", 
    sizes, "px;margin-top: 4px;line-height: ", 
    sizes, "px;'>", 
    labels, "</div>")
  
    return(addLegend(map, position = "bottomright", colors = colorAdditions, labels = labelAdditions, opacity = opacity, title = title))
    
}
```

### Map of Eviction Rates and % ESL speakers, by census tract

Census tracts with a higher percentage of ESL speakers appear to have lower eviction rates. English language nativity appears to be geographically clustered.

```{r ESL, echo=FALSE}
leaflet::leaflet(data = census_tracts_crs) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addPolygons(
    fillColor = ~eng_palette(pct_eng),
    fillOpacity = 0.8,
    color = "BDBDC3",
    weight = 1,
    popup = esl_popup,
    highlightOptions = highlightOptions(color = "black", weight = 2, bringToFront = TRUE)) %>% 
      addLegend('bottomleft',
            pal = eng_palette,
            values = ~pct_eng,
            title = '% ESL',
            opacity = 1) %>%
  addCircleMarkers(
    lat = ~ census_tracts_crs$lat,
    lng = ~ census_tracts_crs$long,
    color = "red",
    radius = (sqrt(census_tracts_crs$eviction_rate)*2),
    weight = 0
    ) %>%
      addLegendCustom(colors = c("red", "red", "red"), 
        labels = c("1", "3", "8"), sizes = c(4, 6.9282, 11.3137))
```

### Map of Eviction Rates and % non-White, by census tract

Census tracts with a higher percentage of non-White residents appear to have higher eviction rates. Racial composition of census tracts appears to be geographically clustered.

```{r non-White, echo=FALSE}
leaflet::leaflet(data = census_tracts_crs) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addPolygons(
    fillColor = ~nonwhite_palette(pct_nonwhite_racedata),
    fillOpacity = 0.8,
    color = "BDBDC3",
    weight = 1,
    popup = nonwhite_popup,
    highlightOptions = highlightOptions(color = "black", weight = 2, bringToFront = TRUE)) %>% 
      addLegend('bottomleft',
            pal = nonwhite_palette,
            values = ~pct_nonwhite_racedata,
            title = '% non-White',
            opacity = 1) %>%
  addCircleMarkers(
    lat = ~ census_tracts_crs$lat,
    lng = ~ census_tracts_crs$long,
    color = "red",
    radius = (sqrt(census_tracts_crs$eviction_rate)*2),
    weight = 0
    ) %>%
      addLegendCustom(colors = c("red", "red", "red"), 
        labels = c("1", "3", "8"), sizes = c(4, 6.9282, 11.3137))
```


### Map of Eviction Rates and Rent Burden, by census tract

The relationship between eviction rates and rent burden were not immediately obvious. However, census tracts with high rent burdens appeared to be geographically clustered.

```{r rent-burden, echo=FALSE}
leaflet::leaflet(data = census_tracts_crs) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addPolygons(
    fillColor = ~rentburden_palette(rent_burden),
    fillOpacity = 0.8,
    color = "BDBDC3",
    weight = 1,
    popup = rentburden_popup,
    highlightOptions = highlightOptions(color = "black", weight = 2, bringToFront = TRUE)) %>% 
      addLegend('bottomleft',
            pal = rentburden_palette,
            values = ~rent_burden,
            title = 'Rent Burden',
            opacity = 1) %>%
  addCircleMarkers(
    lat = ~ census_tracts_crs$lat,
    lng = ~ census_tracts_crs$long,
    color = "red",
    radius = (sqrt(census_tracts_crs$eviction_rate)*2),
    weight = 0
    ) %>%
      addLegendCustom(colors = c("red", "red", "red"), 
        labels = c("1", "3", "8"), sizes = c(4, 6.9282, 11.3137))
```

### Map of Eviction Rates and Population Density, by census tract

A weak relationship seems to exist between population density and eviction rates, with eviction rates lower in more densely populated census tracts. Census tracts with high rent burdens appeared to be weakly geographically clustered.

```{r popdensity, echo=FALSE}
leaflet::leaflet(data = census_tracts_crs) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addPolygons(
    fillColor = ~popdensity_palette(total_pop_densitydata),
    fillOpacity = 0.8,
    color = "BDBDC3",
    weight = 1,
    popup = popdensity_palette,
    highlightOptions = highlightOptions(color = "black", weight = 2, bringToFront = TRUE)) %>% 
      addLegend('bottomleft',
            pal = popdensity_palette,
            values = ~total_pop_densitydata,
            title = 'Population Density',
            opacity = 1) %>%
  addCircleMarkers(
    lat = ~ census_tracts_crs$lat,
    lng = ~ census_tracts_crs$long,
    color = "red",
    radius = (sqrt(census_tracts_crs$eviction_rate)*2),
    weight = 0
    ) %>%
      addLegendCustom(colors = c("red", "red", "red"), 
        labels = c("1", "3", "8"), sizes = c(4, 6.9282, 11.3137))
```

### Findings

Based on our exploratory data visualizations, we identifed significant geographical clustering of our outcome of interest, eviction rates, as well as our predictors of interest (percentage of ESL speakers, percentage of non-White residents by census tract, rent burden, and population density).

With regards to the relationship between our hypothesized predictors and eviction rates, we identified a clear positive association between the percentage of non-White residents and evictions rates in a given census tract. The percentage of ESL speakers in a given census tract appeared to be inversely related to eviction rates; we decided to further examine whether the percentage of non-White residents confounded the relationship between percentage of ESL speakers and eviction rates. Relationships between rent burden, population density, and eviction rates were less obvious; we conducted further exploratory analyses during model building.

