---
title: "Exploring Brooklyn Neighborhoods"
output: html_document
---

<style type="text/css">

h1.title {
  text-align: center;
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(plotly)
library(maps)
library(devtools)
library(leaflet)
library(patchwork)
library(BAMMtools)
library(leafsync)
```


```{r echo = FALSE}
# include = FALSE not working; using work around
suppressPackageStartupMessages(library(rgdal))

```


```{r echo = FALSE, message = FALSE, warning = FALSE}
# include = FALSE not working; using work around
library(maptools)
```


```{r echo = FALSE, message = FALSE}
# loading NYC Open Data file containing neighborhood names by census tract
neighborhood_df = 
  read_excel("./data/NYC_neighborhoods_by_censustract_2010.xlsx", 
             skip = 5, 
             col_names = c("borough", "county_code", "boro_code", "census_tract", "PUMA", "nta_code", "neighborhood")
             ) %>% 
#creating 11 digit FIPS code by pasting country code (36) with `county_code` and `census_tract`
  mutate(geo_id = paste0("36", county_code, census_tract)
  )
```


```{r echo = FALSE, message = FALSE}
#loading/cleaning Eviction Data for the neighborhood maps
eviction_data = read.csv('./data/EvictionData_NY.csv') %>% 
  janitor::clean_names() %>% 
  filter(year == 2016) %>% 
  mutate(
    geoid = as.character(geoid)) %>% 
  filter(parent_location == "Kings County, New York") %>%
  left_join(neighborhood_df, by = c("geoid" = "geo_id"))

neighborhood_data_prep = eviction_data %>% 
  group_by(neighborhood) %>% 
  mutate(
    neighb_pop = sum(population),
    tract_weight = population/neighb_pop,
    eviction_discrepancy = eviction_filing_rate - eviction_rate,
    weighted_eviction_discrep = eviction_discrepancy*tract_weight,
    weighted_eviction_rate = eviction_rate*tract_weight,
    weighted_eviction_filing_rate = eviction_filing_rate*tract_weight,
    pct_nonwhite = 100 - pct_white,
    weighted_pct_nonwhite = pct_nonwhite*tract_weight,
    weighted_povertyrate = poverty_rate*tract_weight)

#creating weighted neighborhood-level variables
neighborhood_data = neighborhood_data_prep %>% 
  summarize(
    wt_eviction_rate = round(as.numeric(sum(weighted_eviction_rate, na.rm = TRUE)), digits = 2),
    wt_eviction_rate_discrep = round(sum(weighted_eviction_discrep, na.rm = TRUE), digits = 2),
    wt_eviction_filing_rate = round(sum(weighted_eviction_filing_rate, na.rm = TRUE), digits = 2),
    wt_pct_nonwhite = round(sum(weighted_pct_nonwhite, na.rm = TRUE), digits = 2),
    wt_povertyrate = round(sum(weighted_povertyrate, na.rm = TRUE), digits = 2)) %>% 
  ungroup() %>% 
  mutate(neighborhood = as.factor(neighborhood))

#updating shapefile of parks, cemeteries, etc. to 'NA'
neighborhood_data <- within(neighborhood_data, wt_eviction_rate[neighborhood == 'park-cemetery-etc-Brooklyn'] <- NA)
neighborhood_data <- within(neighborhood_data, wt_eviction_rate_discrep[neighborhood == 'park-cemetery-etc-Brooklyn'] <- NA)
```


```{r include=FALSE, warning = FALSE, message = FALSE}
#loading NYC neighborhoood shapefile
neighborhood_map = readOGR(dsn = "maps/2010-NTAs.shp", encoding = "UTF-8")

#limiting NYC shapefile to Brooklyn
neighborhood_map = 
  neighborhood_map[neighborhood_map$boro_name == "Brooklyn", ]

writeOGR(neighborhood_map,"./mapping_files","neighborhood_temp", driver = "ESRI Shapefile", overwrite_layer = TRUE)
neighborhood_map = readOGR("./mapping_files","neighborhood_temp")

neighborhood_map@data = left_join(neighborhood_map@data, neighborhood_data, by = c("ntaname" = "neighborhood"))

neighborhood_map_crs = spTransform(neighborhood_map, CRS("+init=epsg:4326"))
```


```{r include=FALSE}
#Eviction rate map set-up

#pop up label
label_popup1 = paste0(
  "<strong>Neighborhood: </strong>",
  neighborhood_map$ntaname,
  "<br><strong>Eviction rate: </strong>",
  neighborhood_map$wt_eviction_rate
)

# get jenks natural breaks
getJenksBreaks(neighborhood_map$wt_eviction_rate, 5)

#set bins based on natural breaks
eviction_rate_bins = c(0.22, 1.04, 1.80, 2.69, 3.57)

#set color palette
eviction_palette1 = colorBin('Reds', bins = eviction_rate_bins, na.color = '#d9d9d9')
```


```{r echo = FALSE, message = FALSE}
#create map of eviction rates
map_evic = leaflet::leaflet(data = neighborhood_map_crs) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addPolygons(
    fillColor = ~eviction_palette1(wt_eviction_rate),
    fillOpacity = 0.8,
    color = "BDBDC3",
    weight = 1,
    popup = label_popup1,
    highlightOptions = highlightOptions(color = "black", weight = 2, bringToFront = TRUE)) %>% 
      addLegend('topright',
                pal = eviction_palette1,
                values = ~wt_eviction_rate,
                title = 'Eviction Rate',
                opacity = 1)
```


```{r include=FALSE}
#discrepancy map set-up

#pop up label
label_popup2 = paste0(
  "<strong>Neighborhood: </strong>",
  neighborhood_map$ntaname,
  "<br><strong>Eviction filing discrepancy rate: </strong>",
  neighborhood_map$wt_eviction_rate_discrep
)

# get jenks natural breaks
getJenksBreaks(neighborhood_map$wt_eviction_rate_discrep, 5)

#set bins based on natural breaks
eviction_rate_discrep_bins = c(0.95, 3.94, 7.23, 11.90, 29.37)

#set color palette
eviction_palette2 = colorBin('Reds', bins = eviction_rate_discrep_bins, na.color = '#d9d9d9')
```


```{r echo = FALSE, message = FALSE}
#create map of discrepancy rates
map_discrep = leaflet::leaflet(data = neighborhood_map_crs) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addPolygons(
    fillColor = ~eviction_palette2(wt_eviction_rate_discrep),
    fillOpacity = 0.8,
    color = "BDBDC3",
    weight = 1,
    popup = label_popup2,
    highlightOptions = highlightOptions(color = "black", weight = 2, bringToFront = TRUE)) %>% 
      addLegend('topright',
                pal = eviction_palette2,
                values = ~wt_eviction_rate_discrep,
                title = 'Rate Discrepancy',
                opacity = 1)

```

<br>

### Mapping evictions and filing discrepancies

There were `r sum(pull(eviction_data, eviction_filings))` eviction filings and `r sum(pull(eviction_data, evictions))` evictions across Brooklyn in 2016. Eviction rates and eviction filing rates represent the number of evictions or eviction filings per neighborhood divided by the number of renter occupied households in a given neighborhood - that value is then multipled by 100 to represent a percentage. `Rate Discrepancy` represents the difference between eviction filing rates and eviction rates. Our motivation for this exploration was the idea that discriminatory practices would manifest in high rate discrepancies, where landlords illegitimately file to evict a tenant. 

```{r echo = FALSE}
#putting maps side by side
sync(map_discrep, map_evic, ncol = 2, sync = "all", sync.cursor = TRUE, no.initial.sync = TRUE)

```


#### __________________________________________________________________________________________________

<br>

### Fifteen neighborhoods with highest variable rates

To further compare eviction rates and filing discrepancies with neighborhood poverty rates and racial composition, we looked at the 10 neighborhoods with the highest rates of poverty, percent nonwhite population, eviction rate, and discrepancy between evicition filing rate and eviction rate. Five neighborhoods appear in all four charts (in red). These findings suggest that eviction filing discrepancies occur in areas that experience other domains of disadvantage. Furthermore, the high filing discrepancy rates may indicate that landlords engage in discriminatory practices against certain populations. 


```{r echo = FALSE, message = FALSE}
#plotly bar chart of eviction rates
barchart_evictions = neighborhood_data %>%
  top_n(15, wt_eviction_rate) %>% 
  mutate(neighborhood = fct_reorder(neighborhood, wt_eviction_rate)) %>% 
  plot_ly(x = ~wt_eviction_rate, y = ~neighborhood, type = 'bar', orientation = 'h', width = 900, height = 700, marker = list(color = c('rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)'))) %>% 
  layout(title = "",
         xaxis = list(title = "Eviction Rate"),
         yaxis = list(title = ""),
         showlegend = FALSE)


#plotly bar chart of eviction filing/rate discrepancies
bar_chart_eviction_discrep = neighborhood_data %>%
  top_n(15, wt_eviction_rate_discrep) %>% 
  mutate(neighborhood = fct_reorder(neighborhood, wt_eviction_rate_discrep)) %>% 
  plot_ly(x = ~wt_eviction_rate_discrep, y = ~neighborhood, type = 'bar', orientation = 'h', width = 1000, height = 700, marker = list(color = c('rgba(222,45,38,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)'))) %>% 
  layout(title = "",
         xaxis = list(title = "Filing Discrepancy"),
         yaxis = list(title = ""),
         showlegend = FALSE)


#plotly bar chart of pct nonwhite
bar_chart_pct_nonwhite = neighborhood_data %>%
  top_n(15, wt_pct_nonwhite) %>% 
  mutate(neighborhood = fct_reorder(neighborhood, wt_pct_nonwhite)) %>% 
  plot_ly(x = ~wt_pct_nonwhite, y = ~neighborhood, type = 'bar', orientation = 'h', width = 900, height = 700, marker = list(color = c('rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)'))) %>% 
  layout(title = "",
         xaxis = list(title = "Percent Nonwhite"),
         yaxis = list(title = ""),
         showlegend = FALSE)


#plotly bar chart of poverty rates
bar_chart_poverty = neighborhood_data %>%
  top_n(15, wt_povertyrate) %>% 
  mutate(neighborhood = fct_reorder(neighborhood, wt_povertyrate)) %>% 
  plot_ly(x = ~wt_povertyrate, y = ~neighborhood, type = 'bar', orientation = 'h', width = 1000, height = 700, marker = list(color = c('rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(222,45,38,1)', 'rgba(222,45,38,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)'))) %>% 
  layout(title = "",
         xaxis = list(title = "Poverty Rate"),
         yaxis = list(title = ""),
         showlegend = FALSE)
  

#combining plots visually
subplot(bar_chart_eviction_discrep,
        barchart_evictions,
        bar_chart_poverty,
        bar_chart_pct_nonwhite,
        nrows = 2, widths = c(0.5, 0.5),
        margin = c(0.1, 0.2, 0.1, 0.1),
        shareX = FALSE, shareY = FALSE,
        titleX = TRUE, titleY = FALSE)

```
