---
title: "Exploratory Spatial Analyses"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(plotly)
library(rgdal)
library(maps)
library(devtools)
library(leaflet)
library(maptools)
library(BAMMtools)
library(patchwork)
```

### Mapping the hypothesized predictors of eviction rates

In order to inform model building, we first observed simple visualizations of our four predictors (percentage of ESL speakers, percentage of non-White residents, rent burden, and population density) over a choropleth of eviction rates in Brooklyn by census tract in 2010.

```{r include = FALSE}
# Loading in Brooklyn data
joined_data_bklyn_2010 = 
  readRDS('./data/joined_data_bklyn') %>%
  filter(year == "2010")

# Loading census tract shapefiles from NYC OpenData:
census_tracts = readOGR(dsn = "maps/2010-census-tracts.shp", encoding = "UTF-8") 

census_tracts = 
  census_tracts[census_tracts$boro_name == "Brooklyn", ]

writeOGR(census_tracts,"./mapping_files","census_temp", driver = "ESRI Shapefile", overwrite_layer = TRUE)
census_tracts = readOGR("./mapping_files","census_temp")

# Constructing geoid for merging
census_tracts@data = 
  census_tracts@data %>% 
  mutate(
    county_code = as.character(recode(boro_code, "1" = "061", "2" = "005", "3" = "047", "4" = "081", "5" = "085")),
    ct2010 = as.character(ct2010),
    fips_11dig = paste0("36", county_code, ct2010),
    long = coordinates(spTransform(census_tracts, CRS("+proj=longlat +datum=WGS84")))[,1],
    lat = coordinates(spTransform(census_tracts, CRS("+proj=longlat +datum=WGS84")))[,2]
    ) %>%
  unnest(c(long, lat))

# Joining data and shapefiles
census_tracts@data = left_join(census_tracts@data, joined_data_bklyn_2010, by = c("fips_11dig" = "geoid"))
```

```{r echo=FALSE, include = FALSE}
census_tracts_crs = spTransform(census_tracts, CRS("+init=epsg:4326"))

# writeOGR(census_tracts_crs, './mapping_files/census_tracts_geojson', layer = 'census_tracts', driver = 'GeoJSON')
```

```{r echo=FALSE, include = FALSE}
#Labels
esl_popup = paste0(
  "<strong>Census tract: </strong>",
  census_tracts_crs$ctlabel,
  "<br><strong>Eviction rate: </strong>",
  census_tracts_crs$eviction_rate,
  "<br><strong>% ESL: </strong>",
  census_tracts_crs$pct_eng
)

nonwhite_popup = paste0(
  "<strong>Census tract: </strong>",
  census_tracts_crs$ctlabel,
  "<br><strong>Eviction rate: </strong>",
  census_tracts_crs$eviction_rate,
  "<br><strong>% non-White: </strong>",
  census_tracts_crs$pct_nonwhite_racedata
)

rentburden_popup = paste0(
  "<strong>Census tract: </strong>",
  census_tracts_crs$ctlabel,
  "<br><strong>Eviction rate: </strong>",
  census_tracts_crs$eviction_rate,
  "<br><strong>Rent burden: </strong>",
  census_tracts_crs$rent_burden
)

popdensity_popup = paste0(
  "<strong>Census tract: </strong>",
  census_tracts_crs$ctlabel,
  "<br><strong>Eviction rate: </strong>",
  census_tracts_crs$eviction_rate,
  "<br><strong>Population Density: </strong>",
  census_tracts_crs$total_pop_densitydata
)

# get jenks natural breaks
getJenksBreaks(census_tracts_crs$eviction_rate, 5)
getJenksBreaks(census_tracts_crs$pct_eng, 5)
getJenksBreaks(census_tracts_crs$pct_nonwhite_racedata, 5)
getJenksBreaks(census_tracts_crs$rent_burden, 5)
getJenksBreaks(census_tracts_crs$total_pop_densitydata, 5)

#set bins based on natural breaks
eviction_rate_bins = c(0, 1.23, 2.51, 4.12, 7.62)
pct_eng_bins = c(0, 14.6, 29.8, 46.9, 82.5)
nonwhite_bins = c(0, 22.9, 49.6, 77.2, 100)
rentburden_bins = c(13.9, 28.4, 34.6, 42.5, 50.0)
popdensity_bins = c(22, 2432, 3636, 5124, 8475)

#set color palette
eviction_palette = colorBin('Blues', bins = eviction_rate_bins, na.color = '#d9d9d9')
eng_palette = colorBin('Reds', bins = pct_eng_bins, na.color = '#d9d9d9')
nonwhite_palette = colorBin('Reds', bins = nonwhite_bins, na.color = '#d9d9d9')
rentburden_palette = colorBin('Reds', bins = rentburden_bins, na.color = '#d9d9d9')
popdensity_palette = colorBin('Reds', bins = popdensity_bins, na.color = '#d9d9d9')
```

Results were as follows: 

### Map of Eviction Rates and % ESL speakers, by census tract

Census tracts with a higher percentage of ESL speakers appear to have lower eviction rates. English language nativity appears to be geographically clustered.

```{r echo=FALSE}
leaflet::leaflet(data = census_tracts_crs) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addPolygons(
    fillColor = ~eviction_palette(eviction_rate),
    fillOpacity = 0.8,
    color = "BDBDC3",
    weight = 1,
    popup = esl_popup,
    highlightOptions = highlightOptions(color = "black", weight = 2, bringToFront = TRUE)) %>% 
      addLegend('bottomleft',
            pal = eviction_palette,
            values = ~eviction_rate,
            title = 'Eviction rate',
            opacity = 1) %>%
  addCircleMarkers(
    lat = ~ census_tracts_crs$lat,
    lng = ~ census_tracts_crs$long,
    color = ~ eng_palette(census_tracts_crs$pct_eng),
    radius = ((census_tracts_crs$pct_eng)^2)/800,
    weight = 1
    ) %>%
      addLegend('bottomright',
            pal = eng_palette,
            values = ~pct_eng,
            title = '% ESL',
            opacity = 1)
```

### Map of Eviction Rates and % non-White, by census tract

Census tracts with a higher percentage of non-White residents appear to have higher eviction rates. Racial composition of census tracts appears to be geographically clustered.

```{r echo=FALSE}
leaflet::leaflet(data = census_tracts_crs) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addPolygons(
    fillColor = ~eviction_palette(eviction_rate),
    fillOpacity = 0.8,
    color = "BDBDC3",
    weight = 1,
    popup = nonwhite_popup,
    highlightOptions = highlightOptions(color = "black", weight = 2, bringToFront = TRUE)) %>% 
      addLegend('bottomleft',
            pal = eviction_palette,
            values = ~eviction_rate,
            title = 'Eviction rate',
            opacity = 1) %>%
  addCircleMarkers(
    lat = ~ census_tracts_crs$lat,
    lng = ~ census_tracts_crs$long,
    color = ~ nonwhite_palette(census_tracts_crs$pct_nonwhite_racedata),
    radius = ((census_tracts_crs$pct_nonwhite_racedata)^1.85)/1000,
    weight = 1
    ) %>%
      addLegend('bottomright',
            pal = nonwhite_palette,
            values = ~pct_nonwhite_racedata,
            title = '% non-White',
            opacity = 1)
```


### Map of Eviction Rates and Rent Burden, by census tract

The relationship between eviction rates and rent burden were not immediately obvious. However, census tracts with high rent burdens appeared to be geographically clustered.

```{r echo=FALSE}
leaflet::leaflet(data = census_tracts_crs) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addPolygons(
    fillColor = ~eviction_palette(eviction_rate),
    fillOpacity = 0.8,
    color = "BDBDC3",
    weight = 1,
    popup = rentburden_popup,
    highlightOptions = highlightOptions(color = "black", weight = 2, bringToFront = TRUE)) %>% 
      addLegend('bottomleft',
            pal = eviction_palette,
            values = ~eviction_rate,
            title = 'Eviction rate',
            opacity = 1) %>%
  addCircleMarkers(
    lat = ~ census_tracts_crs$lat,
    lng = ~ census_tracts_crs$long,
    color = ~ rentburden_palette(census_tracts_crs$rent_burden),
    radius = ((census_tracts_crs$rent_burden)^2.3)/1200,
    weight = 1
    ) %>%
      addLegend('bottomright',
            pal = rentburden_palette,
            values = ~rent_burden,
            title = 'Rent Burden (%)',
            opacity = 1)
```

### Map of Eviction Rates and Population Density, by census tract

A weak relationship seems to exist between population density and eviction rates, with eviction rates lower in more densely populated census tracts. Census tracts with high rent burdens appeared to be weakly geographically clustered.

```{r echo=FALSE}
leaflet::leaflet(data = census_tracts_crs) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addPolygons(
    fillColor = ~eviction_palette(eviction_rate),
    fillOpacity = 0.8,
    color = "BDBDC3",
    weight = 1,
    popup = rentburden_popup,
    highlightOptions = highlightOptions(color = "black", weight = 2, bringToFront = TRUE)) %>% 
      addLegend('bottomleft',
            pal = eviction_palette,
            values = ~eviction_rate,
            title = 'Eviction rate',
            opacity = 1) %>%
  addCircleMarkers(
    lat = ~ census_tracts_crs$lat,
    lng = ~ census_tracts_crs$long,
    color = ~ popdensity_palette(census_tracts_crs$total_pop_densitydata),
    radius = (census_tracts_crs$total_pop_densitydata)^1.3/13000,
    weight = 1
    ) %>%
      addLegend('bottomright',
            pal = popdensity_palette,
            values = ~total_pop_densitydata,
            title = 'Population Density',
            opacity = 1)
```

### Findings

Based on 2010 data on hypothesized predictors of eviction rates in Brooklyn, we identified clear associations between the percentage of non-White residents and evictions rates in a given census tract.

