---
title: "Brooklyn Neighborhoods"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(plotly)
library(maps)
library(devtools)
library(leaflet)
library(patchwork)
library(BAMMtools)
library(leafsync)
```


```{r echo = FALSE}
# include = FALSE not working; using work around
suppressPackageStartupMessages(library(rgdal))

```


```{r echo = FALSE, message = FALSE, warning = FALSE}
library(maptools)
```


```{r echo = FALSE, message = FALSE}
neighborhood_df = 
  read_excel("./data/NYC_neighborhoods_by_censustract_2010.xlsx", 
             skip = 5, 
             col_names = c("borough", "county_code", "boro_code", "census_tract", "PUMA", "nta_code", "neighborhood")
             ) %>% 
#creating 11 digit FIPS code by pasting country code (36) with `county_code` and `census_tract`
  mutate(geo_id = paste0("36", county_code, census_tract)
  )
```


```{r echo = FALSE, message = FALSE}
eviction_data = read.csv('./data/EvictionData_NY.csv') %>% 
  janitor::clean_names() %>% 
  filter(year == 2016) %>% 
  mutate(
    geoid = as.character(geoid)) %>% 
  filter(parent_location == "Kings County, New York") %>% 
  left_join(neighborhood_df, by = c("geoid" = "geo_id"))

neighborhood_data_prep = eviction_data %>% 
  group_by(neighborhood) %>% 
  mutate(
    neighb_pop = sum(population),
    tract_weight = population/neighb_pop,
    eviction_discrepancy = eviction_filing_rate - eviction_rate,
    weighted_eviction_discrep = eviction_discrepancy*tract_weight,
    weighted_eviction_rate = eviction_rate*tract_weight,
    pct_nonwhite = 100 - pct_white,
    weighted_pct_nonwhite = pct_nonwhite*tract_weight,
    weighted_povertyrate = poverty_rate*tract_weight)

neighborhood_data = neighborhood_data_prep %>% 
  summarize(
    wt_eviction_rate = round(as.numeric(sum(weighted_eviction_rate)), digits = 2),
    wt_eviction_rate_discrep = round(sum(weighted_eviction_discrep), digits = 2),
    wt_pct_nonwhite = sum(weighted_pct_nonwhite),
    wt_povertyrate = sum(weighted_povertyrate)) %>% 
  ungroup() %>% 
  mutate(neighborhood = as.factor(neighborhood))

neighborhood_data <- within(neighborhood_data, wt_eviction_rate[neighborhood == 'park-cemetery-etc-Brooklyn'] <- NA)
neighborhood_data <- within(neighborhood_data, wt_eviction_rate_discrep[neighborhood == 'park-cemetery-etc-Brooklyn'] <- NA)
```


```{r include=FALSE, warning = FALSE, message = FALSE}
neighborhood_map = readOGR(dsn = "maps/2010-NTAs.shp", encoding = "UTF-8")

neighborhood_map = 
  neighborhood_map[neighborhood_map$boro_name == "Brooklyn", ]

writeOGR(neighborhood_map,"./mapping_files","neighborhood_temp", driver = "ESRI Shapefile", overwrite_layer = TRUE)
neighborhood_map = readOGR("./mapping_files","neighborhood_temp")

neighborhood_map@data = left_join(neighborhood_map@data, neighborhood_data, by = c("ntaname" = "neighborhood"))


neighborhood_map_crs = spTransform(neighborhood_map, CRS("+init=epsg:4326"))
```


```{r include=FALSE}
#pop up label
label_popup = paste0(
  "<strong>Neighborhood: </strong>",
  neighborhood_map$ntaname,
  "<br><strong>Eviction rate: </strong>",
  neighborhood_map$wt_eviction_rate
)

# get jenks natural breaks
getJenksBreaks(neighborhood_map$wt_eviction_rate, 5)

#set bins based on natural breaks
eviction_rate_bins = c(0.2, 1.0, 1.8, 2.7, 3.6)

#set color palette
eviction_palette1 = colorBin('Reds', bins = eviction_rate_bins, na.color = '#d9d9d9')
```


```{r echo = FALSE, message = FALSE}
#create map
map_evic = leaflet::leaflet(data = neighborhood_map_crs) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addPolygons(
    fillColor = ~eviction_palette1(wt_eviction_rate),
    fillOpacity = 0.8,
    color = "BDBDC3",
    weight = 1,
    popup = label_popup,
    highlightOptions = highlightOptions(color = "black", weight = 2, bringToFront = TRUE)) %>% 
      addLegend('bottomleft',
                pal = eviction_palette1,
                values = ~wt_eviction_rate,
                title = 'Eviction rate',
                opacity = 1)
```


```{r include=FALSE}
#pop up label
label_popup = paste0(
  "<strong>Neighborhood: </strong>",
  neighborhood_map$ntaname,
  "<br><strong>Eviction filing discrepancy rate: </strong>",
  neighborhood_map$wt_eviction_rate_discrep
)

# get jenks natural breaks
getJenksBreaks(neighborhood_map$wt_eviction_rate_discrep, 5)

#set bins based on natural breaks
eviction_rate_discrep_bins = c(0.95, 3.94, 7.23, 11.90, 29.37)

#set color palette
eviction_palette2 = colorBin('Reds', bins = eviction_rate_discrep_bins, na.color = '#d9d9d9')
```


```{r echo = FALSE, message = FALSE}
#create map
map_discrep = leaflet::leaflet(data = neighborhood_map_crs) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addPolygons(
    fillColor = ~eviction_palette2(wt_eviction_rate_discrep),
    fillOpacity = 0.8,
    color = "BDBDC3",
    weight = 1,
    popup = label_popup,
    highlightOptions = highlightOptions(color = "black", weight = 2, bringToFront = TRUE)) %>% 
      addLegend('bottomleft',
                pal = eviction_palette2,
                values = ~wt_eviction_rate_discrep,
                title = 'Eviction filing/eviction rate discrepancy',
                opacity = 1)

```


<br>

There were `r sum(pull(eviction_data, eviction_filings))` eviction filings and `r sum(pull(eviction_data, evictions))` evictions across Brooklyn in 2016. Eviction rates and eviction filing rates represent the number of evictions or eviction filings per neighborhood divided by the number of renter occupied households in a given neighborhood - that value is then multipled by 100 to represent a percentage. 

```{r echo = FALSE}
sync(map_discrep, map_evic, sync = "all", sync.cursor = TRUE, no.initial.sync = TRUE)  

#creating space  
```

<br><br><br>


```{r echo = FALSE, message = FALSE}
#bar chart of evcition rate
barchart_evictions = neighborhood_data %>%
  top_n(15, wt_eviction_rate) %>% 
  mutate(
    neighborhood = fct_reorder(neighborhood, wt_eviction_rate),
    highlight = ifelse(neighborhood == "Brownsville" | neighborhood == "East New York (Pennsylvania Ave)" | neighborhood == "Ocean Hill" | neighborhood == "East New York" | neighborhood == "Crown Heights North", "yes", "no")) %>% 
  ggplot(aes(x = neighborhood, y = wt_eviction_rate, fill = highlight)) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values = c("yes" = "red", "no" = "gray"), guide = FALSE) +
    coord_flip() + 
    labs(y = "Eviction Rate", x = "", title = "Brooklyn Neighborhood Characteristics, 2016")


#bar chart of eviction filing/rate discrepancies
bar_chart_eviction_discrep = neighborhood_data %>%
  top_n(15, wt_eviction_rate_discrep) %>% 
  mutate(
    neighborhood = fct_reorder(neighborhood, wt_eviction_rate_discrep),
    highlight = ifelse(neighborhood == "Brownsville" | neighborhood == "East New York (Pennsylvania Ave)" | neighborhood == "Ocean Hill" | neighborhood == "East New York" | neighborhood == "Crown Heights North", "yes", "no")) %>% 
  ggplot(aes(x = neighborhood, y = wt_eviction_rate_discrep, fill = highlight)) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values = c("yes" = "red", "no" = "gray"), guide = FALSE) +
    coord_flip() +
    labs(y = "Eviction Filing/Rate Discrepancy", x = "")


#bar chart of pct nonwhite
bar_chart_pct_nonwhite = neighborhood_data %>%
  top_n(15, wt_pct_nonwhite) %>% 
  mutate(
    neighborhood = fct_reorder(neighborhood, wt_pct_nonwhite),
    highlight = ifelse(neighborhood == "Brownsville" | neighborhood == "East New York (Pennsylvania Ave)" | neighborhood == "Ocean Hill" | neighborhood == "East New York" | neighborhood == "Crown Heights North", "yes", "no")) %>% 
  ggplot(aes(x = neighborhood, y = wt_pct_nonwhite, fill = highlight)) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values = c("yes" = "red", "no" = "gray"), guide = FALSE) +
    coord_flip() +
    labs(y = "Percent Nonwhite", x = "")


#bar chart of poverty rates
bar_chart_poverty = neighborhood_data %>%
  top_n(15, wt_povertyrate) %>% 
  mutate(
    neighborhood = fct_reorder(neighborhood, wt_povertyrate),
    highlight = ifelse(neighborhood == "Brownsville" | neighborhood == "East New York (Pennsylvania Ave)" | neighborhood == "Ocean Hill" | neighborhood == "East New York" | neighborhood == "Crown Heights North", "yes", "no")) %>% 
  ggplot(aes(x = neighborhood, y = wt_povertyrate, fill = highlight)) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values = c("yes" = "red", "no" = "gray"), guide = FALSE) +
    coord_flip() +
    labs(y = "Poverty Rate", x = "")
```

<br>



To further compare eviction rates and filing discrepancies with neighborhood poverty rates and racial composition, we looked at the 10 neighborhoods with the highest rates of poverty, percent nonwhite population, eviction rate, and discrepancy between eviction rate and filing rate. Five neighborhoods appear in all four charts (in red).

<br>

```{r echo = FALSE, message = FALSE}


#(barchart_evictions + bar_chart_eviction_discrep) / (bar_chart_pct_nonwhite + bar_chart_poverty)

```



```{r echo = FALSE, message = FALSE}

#plotly bar chart of evcition rate
barchart_evictions2 = neighborhood_data %>%
  top_n(15, wt_eviction_rate) %>% 
  mutate(neighborhood = fct_reorder(neighborhood, wt_eviction_rate)) %>% 
  plot_ly(x = ~wt_eviction_rate, y = ~neighborhood, type = 'bar', orientation = 'h', width = 1000, height = 700, marker = list(color = c('rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)'))) %>% 
  layout(title = "",
         xaxis = list(title = "Eviction Rate"),
         yaxis = list(title = ""),
         showlegend = FALSE)


#bar chart of eviction filing/rate discrepancies
bar_chart_eviction_discrep2 = neighborhood_data %>%
  top_n(15, wt_eviction_rate_discrep) %>% 
  mutate(neighborhood = fct_reorder(neighborhood, wt_eviction_rate_discrep)) %>% 
  plot_ly(x = ~wt_eviction_rate_discrep, y = ~neighborhood, type = 'bar', orientation = 'h', width = 1000, height = 700, marker = list(color = c('rgba(222,45,38,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)'))) %>% 
  layout(title = "",
         xaxis = list(title = "Filing Discrepancy"),
         yaxis = list(title = ""),
         showlegend = FALSE)


#plotly bar chart of pct nonwhite
bar_chart_pct_nonwhite2 = neighborhood_data %>%
  top_n(15, wt_pct_nonwhite) %>% 
  mutate(neighborhood = fct_reorder(neighborhood, wt_pct_nonwhite)) %>% 
  plot_ly(x = ~wt_pct_nonwhite, y = ~neighborhood, type = 'bar', orientation = 'h', width = 1000, height = 700, marker = list(color = c('rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)'))) %>% 
  layout(title = "",
         xaxis = list(title = "Percent Nonwhite"),
         yaxis = list(title = ""),
         showlegend = FALSE)


#plotly bar chart of poverty rates
bar_chart_poverty2 = neighborhood_data %>%
  top_n(15, wt_povertyrate) %>% 
  mutate(neighborhood = fct_reorder(neighborhood, wt_povertyrate)) %>% 
  plot_ly(x = ~wt_povertyrate, y = ~neighborhood, type = 'bar', orientation = 'h', width = 1000, height = 700, marker = list(color = c('rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(222,45,38,1)', 'rgba(222,45,38,1)', 'rgba(222,45,38,1)', 'rgba(222,45,38,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)', 'rgba(204,204,204,1)'))) %>% 
  layout(title = "",
         xaxis = list(title = "Poverty Rate"),
         yaxis = list(title = ""),
         showlegend = FALSE)
  


subplot(bar_chart_eviction_discrep2,
        barchart_evictions2,
        bar_chart_poverty2,
        bar_chart_pct_nonwhite2,
        nrows = 2, widths = c(0.5, 0.5),
        margin = c(0.1, 0.2, 0.05, 0.05),
        shareX = FALSE, shareY = FALSE,
        titleX = TRUE, titleY = FALSE)

```
